\propertitle

### 2019-12-07 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-06T17:16:23+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Answering this kind of questions always engages an element of uncertainty.
We cannot say \enquote{the answer is such and such}; at best we can assign
degrees of reasonable belief -- that is, probabilities -- to the possible
answers. The assessment of such distributions of belief involves
experimental data, such as recordings of neuronal activity from specific
brain areas, and pre-data knowledge and hypotheses about biological
conditions and mechanisms. When we translate such pre-data beliefs into an
initial probability distribution we often simplify them in more or less
realistic ways, to make the analysis mathematically tractable. This is why
such initial probability distributions are called \enquote{models}.

The best experimental measurements of instantaneous neuronal activity use
remarkable technologies, but can still only record a very small sample of
neurons -- hundreds at most -- compared to the numbers that constitute a
functionally distinguished brain region. Many probabilistic models focus on
such samples only: they neglect that the recorded neurons are a sample from
a larger population. Such isolating assumptions sometimes escape attention,
being subtly hidden in the mathematics. Some probabilistic models try to
take unrecorded neurons into account, but they do so by describing each
neuron individually, thus becoming very complex
\citep{huang2015,battistinetal2017}. It would be useful to explore models
that operate in between, addressing the activity of the larger brain area
whence the sample comes, but \emph{collectively}, that is, without asking
about individual neuronal details. Such intermediate models would be useful
at least for preliminary investigations, to help us to decide which
hypotheses to discard and which to consider for more complex and costly
models, or to suggest new hypotheses.

% \mynote{[Yasser:] In the present work we aim to address the issue of
%   building probability distributions over a large population using
%   recording from subpopulations by considering the distribution of total
%   population activity as an example. [} 
In the present work we propose such an intermediate probabilistic model. It
answers this question: How much was the \emph{total} activity of a large
neuronal population, given the observation of the activity of a very small
sample thereof? This model addresses a larger brain area, avoiding the
assumption of isolation of the sample; and by focusing on the total
activity, rather than the activity of individual neurons, it remains simple
and numerically tractable.

@@

We illustrate possible uses of our model in \sect~\ref{sec:application}, by
applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012},
\item the activity of 159 neurons recorded for 15 min from a macaque's
  Motor Cortex \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives us the most plausible frequencies of all
levels of total activity of a much larger population, 1\,000--10\,000
neurons in size, during the recording. For example it can tell us that 60
out of 10\,000 neurons were most likely active during 1\% of the recording
time (though not necessarily always the same 60), 250 neurons out of
10\,000 were active during 0.4\% of the recording time, and so on. The
precise meaning of this frequency distribution is explained in
\sect~\ref{sec:model}. For the two example data sets, the guessed frequency
distributions for the larger population are distinctly different from those
for the sample. For the first data set the frequency distribution for the
larger population has two very distinct modes, both at low
activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
(see \fig~\ref{fig:plots_5mom}), whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. The
frequency distribution for the second data set presents one prominent
shoulder in its low-activity mode. These results show that the proposed
model can lead to the formulation or preliminary assessment of interesting
hypotheses, as we will illustrate with toy examples in
\sects~\ref{sec:application} and~\ref{sec:marginalization}. Note that it is
not possible to guess these features of the larger population by applying
the maximum-entropy method \emph{at the level of the sample alone}.

@@

How large is the full population addressed by the model proposed here? Its
size must have some limit and can't obviously include the full brain. The
size is determined by the validity of the formulae from sampling theory and
discussed in \sect~\ref{sec:N}.

In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
approximations on which the model is based, from the point of view of the
probability calculus.

@@

Our notation and terminology follow \textsc{iso}
\citey{iso1993,iso2006,iso2006b} standards and Jaynes
\citey{jaynes1994_r2003} for \dobs. We use \enquote{\dob},
\enquote{belief}, and \enquote{probability} interchangeably.

@@

Our questions concern general features of the total activity $\yA$ of the
larger population during and after the recording, and across sessions under
the same study conditions. For example: what was its frequency distribution
$\yF$ during the recording? How much does this frequency distribution
change across sessions? How much total activity should we expect at any
time bin during a recording? The model presented here gives a
probability distribution that approximately answers all these questions.

@@

The idea behind our approach is easily summarized:
\begin{enumerate}[label=(\alph*)]
\item\label{item:sample_step} Using sampling theory we determine the
  relation between some expected values -- specifically, moments -- for the
  total activity $\ya$ of the sample and corresponding expected values %$M$
  for the total activity $\yA$ of the larger population.
\item\label{item:maxent_step} Using the maximum-entropy method we build a
  distribution $\px(\yA \| M, N)$ for the total activity of the larger
  population of $N$ neurons, using as constraints the expected values $M$
  found in the previous step.
\end{enumerate}
We now discuss the ideas behind these steps more in detail, but leave their
precise mathematical implementation and a more detailed list of references
to appendix~\ref{sec:derivation_maxent_fullnet}.

@@

Step~\ref{item:sample_step} is just an application of the probability
calculus, which gives an exact linear relation between the first $m$
moments for the larger population and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
marginal $\pxx(\ya_{t})$ for the sample activity at that bin.

This sampling relation is even more straightforward if instead of power
moments we use \emph{normalized factorial moments} \citep{potts1953}. The
$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
\begin{equation}
  \label{eq:def_factorial_moment_a}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  \defd
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya),
  \qquad 1\le m \le n,
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
that is, the expected number of distinct $m$-tuples of simultaneously
active neurons (within a time-bin's width), normalized by the maximum
possible number of distinct $m$-tuples. % For
% example, with $m=2$, if $a=4$ neurons spike in a population of $n=5$, we
% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
% normalized number of spiking pairs is therefore $6/10$.
Note that the first $m$ factorial moments together provide the same
information as the first $m$ power moments together, and vice versa: they
are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
degree $m$. So we'll just say \enquote{first $m$ moments} from now
on. % Specifying one set is therefore
% equivalent to specifying the other set.
But the normalized factorial moments have a mathematically convenient
property for our analysis: \emph{the first $n$ normalized factorial moments
  for the sample and for the larger population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  \begin{split}
  \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  &=
  \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
  \qquad\text{or}\\
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
&=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  1\le m \le n.
\end{split}
\end{equation}
% where $\pxxx(\yA)$ is our distribution of belief about the larger-population
% activity, and $\pxx$

@@

In our case we don't know the time averages of the quantity
$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectation for the larger population are determined by that for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured averages} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{larger-population moments}}_{\textit{sampling theory}}
\end{equation*}
We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
$N$ neurons by constraining some of its factorial moments, for example
$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
averages. In formulae, the constraints on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
\underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
  \mathrel{\;=\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}},
   \quad m\in M.
\end{equation}

The result is the distribution of the form
\begin{equation}
  \label{eq:app_maxent_pop_first}
  \begin{gathered}
  \px(\yA \| M, N)  = \frac{1}{Z(\yll)}\;
%  g(\yA)\;
  \exp\Biggl[\sum_{m}\yl_{m}\;
  \binom{\yA}{m}\bigg/\binom{N}{m}
  \Biggr]
  \\
\text{with}\qquad  Z(\yll)\defd \sum_{\yA}
  % g(\yA)\;
  \exp\Biggl[\sum_{m}\yl_{m}\;
  \binom{\yA}{m}\bigg/\binom{N}{m}
  \Biggr],
\end{gathered}
\end{equation}
where the parameters $\yll$ are determined by the
constraints~\eqref{eq:final_constraint}. This formula is further discussed
in appendix~\ref{sec:derivation_maxent_fullnet}.

The amount and degrees of the constraining moments depend on the questions
and hypotheses that a researcher is exploring. We give some examples in the
next two sections. Note that the constraint
equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
if $N$ is strictly larger than $n$; and the possible discrepancy typically
increases with the number of constraints. This discrepancy comes from the
approximate character of two assumptions behind the maximum-entropy method
and of one assumption specific to our application: that the order of the
time bins is irrelevant, that measured averages equal expected values (or
equivalently, the number of time bins $T$ is infinite), and that an
activity level $\yA$ can equally likely be generated by any set of $\yA$
neurons in the larger population. The magnitude of this discrepancy can be
a signature of the (at least temporary) presence of neuronal
\enquote{assemblies} or \enquote{clusters}
\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
in \sect~\ref{sec:assumptions}.

%, but our main goal is to simply introduce
% the general method without entering into biological questions of this
% kind.

@@

Before applying the formula above to two concrete data sets we want to add
two remarks about the maximum-entropy method that are seldom made in the
neuroscientific literature. They are important for the interpretation of
the results and are further discussed in \sect~\ref{sec:assumptions}.
First, the maximum-entropy
method % (based on the \emph{Shannon} relative
% entropy, as opposed to other entropy functions such as Burg's
% \citeyear{burg1975})
rests on some implicit assumptions about the probabilities for the long-run
frequency distribution of activities, besides the assumptions just
mentioned at the end of the previous paragraph. So the often-heard
statement that it gives \enquote{the maximally unbiased (or non-committal)
  distribution} must be taken with a grain of salt.
\citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second, a
maximum-entropy distribution like $\px(\yA \| M, N)$ is the zeroth-order
approximation (in the sense of Laplace's approximation
\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
four distinct distributions for the larger population, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item the most probable \emph{frequency} distribution for the total
  activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among the \emph{recorded} ones,
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
In the present case the validity of the maximum-entropy approximation
decreases as the ratio $nN/T$ increases (see \sect~\ref{sec:assumptions}).

@@

The crucial point is our belief distribution for the activity of the
sample conditional on the activity of the larger population, given by
formula~\eqref{eq:conditional_hypergeometric} in
appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
leading to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests.
This conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
ensues when our degree of belief that the $i$th unit (ball or neuron) will
be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
consider the pool of all neurons which we believe -- with equal degree for
each neuron -- could have been recorded. $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording was made: if we believe
that it doesn't matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-12-07T14:46:25+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Inferring the total activity of a large neuronal population\\ from a small sample}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

Answering this kind of questions always engages an element of uncertainty.
We cannot say \enquote{the answer is such and such}; at best we can assign
degrees of reasonable belief -- that is, probabilities -- to the possible
answers. The assessment of such distributions of belief involves
experimental data, such as recordings of neuronal activity from specific
brain areas, and pre-data knowledge and hypotheses about biological
conditions and mechanisms. When we translate such pre-data beliefs into an
initial probability distribution we often simplify them in more or less
realistic ways, to make the analysis mathematically tractable. This is why
such initial probability distributions are called \enquote{models}; and
that is how we intend this word in the present work.

The best experimental measurements of instantaneous neuronal activity use
remarkable technologies, but can still only record a very small sample of
neurons -- hundreds at most -- compared to the numbers that constitute a
functionally distinguished brain region. Many probabilistic models focus on
such samples only: they neglect that the recorded neurons are a sample from
a larger population. Such isolating assumptions sometimes escape attention,
being subtly hidden in the mathematics. Some probabilistic models try to
take unrecorded neurons into account, but they do so by describing each
neuron individually, thus becoming very complex
\citep{huang2015,battistinetal2017}. It would be useful to explore models
that operate in between, addressing the activity of the larger brain area
whence the sample comes, but \emph{collectively}, that is, without asking
about individual neuronal details. Such intermediate models would be useful
at least for preliminary investigations, to help us to decide which
hypotheses to consider or discard for more complex and costly models, or to
suggest new hypotheses.

% \mynote{[Yasser:] In the present work we aim to address the issue of
%   building probability distributions over a large population using
%   recording from subpopulations by considering the distribution of total
%   population activity as an example. [} 
In the present work we propose such an intermediate probabilistic model. It
answers this question: how much was the \emph{total} activity of a large
neuronal population, given the observation of the activity of a very small
sample thereof? This model addresses a larger brain area, avoiding the
assumption of isolation of the sample; and by focusing on the total
activity, rather than the activity of individual neurons, it remains simple
and numerically tractable.

@@

We illustrate possible uses of our model in \sect~\ref{sec:application}, by
applying it to two concrete data sets:
\begin{enumerate*}[label=(\alph*)]
\item the activity of 65 neurons recorded for 20 min from a rat's Medial
  Entorhinal Cortex \citep{stensolaetal2012},
\item the activity of 159 neurons recorded for 15 min from a macaque's
  Motor Cortex \citep{rostamietal2016_r2017}.
\end{enumerate*}
For each data set the model gives us the most plausible frequencies of all
levels of total activity of a much larger population, 1\,000--10\,000
neurons in size, during the recording. For example it can tell us that 60
out of 10\,000 neurons were most likely active during 1\% of the recording
time (though not necessarily always the same 60), 250 neurons out of
10\,000 were active during 0.4\% of the recording time, and so on. The
precise meaning of this frequency distribution is explained in
\sect~\ref{sec:model}. For the two example data sets, the guessed frequency
distributions for the larger population are distinctly different from those
for the sample. For example, for the first data set the frequency
distribution for the larger population has two very distinct modes, both at
low activities %($0.7\%$ and $2.6\%$ of simultaneously active neurons
% within $3\;\mathrm{ms}$,
(see \fig~\ref{fig:plots_5mom}), whereas the frequency distribution for the
sample is monotonically decreasing with its maximum at zero activity. These
results show that the proposed model can be used for the formulation or
preliminary assessment of interesting hypotheses. We illustrate this use
with toy examples in \sects~\ref{sec:application}
and~\ref{sec:marginalization}. Note that it is not possible to guess these
features of the larger population by applying the maximum-entropy method
\emph{at the level of the sample alone}.

@@

How large is the \enquote{larger population} considered by the model
proposed here? It obviously cannot include the full brain. The size of the
larger population is determined by the validity of the formulae from
sampling theory, and can range from a set of neurons around the recording
probe to part of a brain area, depending on some assumptions about the
recording procedure and the brain region under consideration. This matter
is discussed in detail in \sect~\ref{sec:N}.

In \sect~\ref{sec:assumptions} we discuss in detail the assumptions and
approximations which define our model, from the point of view of the
probability calculus.

@@

Our notation and terminology follow \textsc{iso}
\citey{iso1993,iso2006,iso2006b} standards and Jaynes
\citey{jaynes1994_r2003} for probability. We use \enquote{\dob},
\enquote{belief}, and \enquote{probability} interchangeably.

@@

Our questions concern general features of the total activity $\yA$ of the
larger population during and after the recording, and across sessions under
the same study conditions. For example: what was its frequency distribution
$\yF$ during the recording? How much does this frequency distribution
change across sessions? How much total activity should we expect at any
time bin during a recording? The model presented here gives a probability
distribution over the answers to these questions.

@@

The idea behind our approach can be easily summarized in terms of the
maximum-entropy method:
\begin{enumerate}[label=(\alph*)]
\item\label{item:sample_step} Using sampling theory we determine the
  relation between some expected values -- specifically, moments -- for the
  total activity $\ya$ of the sample and corresponding expected values %$M$
  for the total activity $\yA$ of the larger population.
\item\label{item:maxent_step} Using the maximum-entropy method we build a
  distribution $\px(\yA \| M, N)$ for the total activity of the larger
  population of $N$ neurons, using as constraints the expected values $M$
  found in the previous step.
\end{enumerate}
We now discuss these steps more in detail, but leave their precise
mathematical implementation and a more detailed list of references to
appendix~\ref{sec:derivation_maxent_fullnet}.

@@

Step~\ref{item:sample_step} is just an application of the probability
calculus, which gives an exact linear relation between the first $m$
moments for the larger population and the first $m$ for the sample
\citep[\eqns~(16)]{portamanaetal2015}. The ones determine the others and
vice versa at every time bin. This relation holds for any belief
distribution $\pxxx(\yA_{t})$ for the larger-population activity and its
marginal $\pxx(\ya_{t})$ for the sample activity at that bin. These two
distributions are related by
\begin{equation}
  \label{eq:conditional_hypergeometric_t}
  \pxx(\ya_{t})= \sum_{\yA_{t}=0}^{N} \yG_{\ya_{t}\,\yA_{t}} \;\pxxx(\yA_{t})
  \qquad \ya_{t} \in \set{0,\dotsc,n},
\end{equation}
where $\yG_{\ya\yA}$ is the hypergeometric distribution defined in
\eqn~\eqref{eq:conditional_hypergeometric}, characteristic of
\enquote{drawing without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.

This sampling relation is even more straightforward if instead of power
moments we use \emph{normalized factorial moments} \citep{potts1953}. The
$m$th normalized factorial moment of a distribution $\pxx(\ya)$ is
\begin{equation}
  \label{eq:def_factorial_moment_a}
  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  % \defd
  \sum_{\ya=0}^{n} %m!\,
  \displaystyle\binom{\ya}{m} \bigg/ \binom{n}{m}\; \pxx(\ya)
  \qquad m \in \set{1,\dotsc,n},
\end{equation}
% where $\px(\ya \| \yH)$ is the probability for activity $\ya$ given
% assumptions $\yH$; \enquote{$\ya$} may refer to any recorded bin or to a
% new bin.
that is, the expected number of distinct $m$-tuples of simultaneously
active neurons (within a time-bin's width), normalized by the maximum
possible number of distinct $m$-tuples. % For
% example, with $m=2$, if $a=4$ neurons spike in a population of $n=5$, we
% have $\tbinom{4}{2}=6$ distinct pairs of simultaneously spiking neurons,
% and the total number of distinct pairs is $\tbinom{5}{2}=10$. The
% normalized number of spiking pairs is therefore $6/10$.
Note that the first $m$ factorial moments together provide the same
information as the first $m$ power moments together, and vice versa: they
are linearly related because $\tbinom{\ya}{m}$ is a polynomial in $\ya$ of
degree $m$. So we'll just say \enquote{first $m$ moments} from now
on. % Specifying one set is therefore
% equivalent to specifying the other set.
But the normalized factorial moments have a mathematically convenient
property: \emph{the first $n$ normalized factorial moments for the sample
  and for the larger population are numerically identical}:
\begin{equation}
  \label{eq:equal_m_factorialmoments}
  % \begin{split}
  % \E\Biggl[\binom{\ya}{m} \Biggr]\bigg/\binom{n}{m}
  % &=
  % \E\Biggl[\binom{\yA}{m} \Biggr]\bigg/\binom{N}{m}
  % \qquad\text{or}\\
    \sum_{\ya=0}^{n} %m!\,
  \binom{\ya}{m}\bigg/  \binom{n}{m}\; \pxx(\ya)
=    \sum_{\yA=0}^{N} %m!\,
  \binom{\yA}{m}\bigg/  \binom{N}{m}\; \pxxx(\yA),\qquad
  % \E\condt[\big]{\tbinom{\ya}{m}\;\tbinom{n}{m}^{-1}}{\yH} =
  % \E\condt[\big]{\tbinom{\yA}{m}\;\tbinom{N}{m}^{-1}}{\yH},
  m \in \set{1,\dotsc,n}.
%\end{split}
\end{equation}
% where $\pxxx(\yA)$ is our distribution of belief about the larger-population
% activity, and $\pxx$

@@

In our case we don't know the time averages of the quantity
$\binom{\yA}{m}$, so we cannot directly equate them to the factorial moment
$\E\bigl[\binom{\yA}{m}\bigr]$ of the larger-population distribution
$\pxxx(\yA)$. But \eqn~\eqref{eq:equal_m_factorialmoments} of
step~\ref{item:sample_step} comes to our rescue, because it says that the
expectation for the larger population are determined by that for the sample
$\E\bigl[\binom{\ya}{m}\bigr]$, and we do have the time average of its
corresponding sample quantity $\binom{\ya}{m}$. So we can combine the two
steps:
\begin{equation*}
  \label{eq:chain_prescription_moments}
  \mathrlap{\overbracket{\phantom{\text{measured averages} \rightarrow
        \text{sample moments}}}^{\textit{maximum-entropy prescription}}}
  \text{measured averages} \rightarrow
        \underbracket{\text{sample moments}
  \rightarrow
    \text{larger-population moments}}_{\textit{sampling theory}}
\end{equation*}
We obtain a distribution $\px(\yA \| M, N)$ for the larger population of
$N$ neurons by constraining some of its factorial moments, for example
$M=\set{1,\dotsc,m'}$ with $m'\le n$, to be equal to the sample's recorded
averages. In formulae, the constraints on $\pxxx(\yA)$ are
\begin{equation}
  \label{eq:final_constraint}
  \underbracket{\frac{1}{T}\sum_{t}\binom{\ya_{t}}{m}\bigg/\binom{n}{m}
  \equiv
   \sum_{\ya}
  \binom{\ya}{m}\bigg/  \binom{n}{m} \; \yff_{\ya}}_{\text{measured averages}}
  \mathrel{\;\;=\;\;}
   \underbracket{\sum_{\yA}
     \binom{\yA}{m}\bigg/  \binom{N}{m} \; \pxxx(\yA)}_{\text{distribution moments}}
   \qquad m\in M.
\end{equation}
\smallskip

The result is the distribution of the form
\begin{equation}
  \label{eq:app_maxent_pop_first}
  \begin{gathered}
  \px(\yA \| M, N)  = \frac{1}{Z(\yll)}\;
%  g(\yA)\;
  \exp\Biggl[\sum_{m}\yl_{m}\;
  \binom{\yA}{m}\bigg/\binom{N}{m}
  \Biggr]
  \\
\text{with}\qquad  Z(\yll)\defd \sum_{\yA}
  % g(\yA)\;
  \exp\Biggl[\sum_{m}\yl_{m}\;
  \binom{\yA}{m}\bigg/\binom{N}{m}
  \Biggr],
\end{gathered}
\end{equation}
where the $m'$ parameters $\yll \defd (\yl_{m})$ are determined by the
constraints~\eqref{eq:final_constraint}. This formula is further discussed
in appendix~\ref{sec:derivation_maxent_fullnet}.

The amount and degrees of the constraining moments depend on the questions
and hypotheses that a researcher is exploring. We give some examples in the
next two sections. Note that the constraint
equation~\eqref{eq:final_constraint} for the
distribution~\eqref{eq:app_maxent_pop_first} may not have exact solutions
if $N$ is strictly larger than $n$. The possible discrepancy between its
two sides typically increases with the number of constraints. This
discrepancy comes from an approximation implicit in the maximum-entropy
method combined with one assumption specific to our application: that
measured averages equal expected values (or equivalently, the number of
time bins $T$ is infinite), and that an activity level $\yA$ can equally
likely be generated by any set of $\yA$ neurons in the larger population.
The magnitude of this discrepancy can be a signature of the presence of
neuronal \enquote{assemblies} or \enquote{clusters}
\citep[\chap~12]{gerstneretal2014}{hebb1949_r2002}. We discuss this matter
in \sect~\ref{sec:assumptions}.

@@

One important question remains about the distribution $\px$ of
\eqn~\eqref{eq:app_maxent_pop_first}: \emph{what} is it a distribution of?
-- Of probability? Of relative frequency? The answer is related to an
important fact about the maximum-entropy method, and also to our
alternative point of view of the present model.

% Before applying the formula above to two concrete data sets we want to add
% two remarks about the maximum-entropy method that are seldom made in the
% neuroscientific literature. They are important for the interpretation of
% the results and are further discussed in \sect~\ref{sec:assumptions}.
% First, the maximum-entropy
% method % (based on the \emph{Shannon} relative
% % entropy, as opposed to other entropy functions such as Burg's
% % \citeyear{burg1975})
% rests on some implicit assumptions about the probabilities for the long-run
% frequency distribution of activities, besides the assumptions just
% mentioned at the end of the previous paragraph. So the often-heard
% statement that it gives \enquote{the maximally unbiased (or non-committal)
%   distribution} must be taken with a grain of salt.
% \citep{jaynes1986d_r1996,portamana2009,portamana2017}. Second,
A maximum-entropy distribution like $\px(\yA \| M, N)$ is equivalent to the
zeroth-order approximation (in the sense of Laplace's approximation
\citep[\chap~4]{debruijn1958_r1961}{tierneyetal1986,strawderman2000}) of
four distinct distributions for the larger population, which differ
numerically from one another in higher-order approximations:
\begin{enumerate}[label=(\roman*)]
\item\label{item:prob_freq} the most probable \emph{frequency} distribution
  for the total activity across the \emph{recorded} bins, %$\yFF(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at any time bin among the \emph{recorded} ones,
  %$\yp(\yA_{t} \| \yH)$ with $1\le t \le T$,
\item the most probable \emph{frequency} distribution for the total
  activity in a very long run of \emph{new} time bins, %$\ynuu(\yA)$,
\item the \emph{belief} distribution for the value of the total
  activity at a \emph{new} time bin.
  % $\yp(\yA_{t} \| \yH)$ with $t<1$ or $t>T$
\end{enumerate}
Their mathematical relationship is presented more explicitly in
appendix~\ref{sec:inference_prob}. The
distribution~\eqref{eq:app_maxent_pop_first} can therefore be interpreted
in each of the four ways above, in this approximate sense. In the present
case the validity of the approximation decreases as the ratio $nN/T$
increases. The most robust interpretation is~\ref{item:prob_freq}; and this
is how we prefer to interpret the distribution $\px$: as \emph{the most
  probable relative-frequency distribution} of the activity $\yA$. See
\sect~\ref{sec:assumptions} for further discussion about assumptions and
approximations.

@@

The crucial point is our belief distribution for the activity of the sample
\emph{conditional} on the activity of the larger population. It is given by
the hypergeometric distribution~\eqref{eq:conditional_hypergeometric},
reprinted here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
    \binom{\yA}{\ya}\binom{N-\yA}{n-\ya}\binom{N}{n}^{-1}
\equiv \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}.
\end{equation}
This distribution leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our model rests. This
conditional probability, characteristic of \enquote{drawing without
  replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968},
ensues when our degree of belief that the $i$th unit (ball or neuron) will
be drawn has the same value for all $i \in \set{1,\dotsc,N}$. Thus,
consider the pool of all neurons which we believe -- with equal degree for
each neuron -- could have been recorded. $N$ is the size of that pool. $N$
therefore depends on factors such as: the shape, dimensions, and technical
specifications of the recording probe; the inaccuracy in the insertion of
the probe, leading for example to slightly different insertion points or
angles; the density of neurons around the probe. But it also depends on the
homogeneity of the brain region where the recording was made: if we believe
that it doesn't matter whether the probe had been inserted in some other
point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. $N$ can
therefore only be assessed case by case.
